---
title: Appendix S1. Instructions for retrieving and archiving the environmental covariates.
subtitle: Supporting information for Scheuerell et al.
output:
  pdf_document:
    highlight: haddock
    toc: yes
    number_sections: true
    toc_depth: '3'
fontsize: 11pt
geometry: margin=1in
---

```{r set_options, echo = FALSE, message = FALSE}
options(width = 100)
knitr::opts_chunk$set(message = FALSE)
```

\vspace{0.2in}

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

# Background

This appendix describes how to retrieve the environmental covariates used in our analyses. After reading in the raw data, summarizing them (if necessary), and trimming them to the appropriate time frame, they table of covariates is written to a `.csv` file.

All of the analyses require the [R software](https://cran.r-project.org/) (v3.4.3 or later) for data retrieval and processing. We also need the __readr__ and __here__ packages, which are not included with the base installation of __R__.

```{r load_pkgs, message = FALSE, warning = FALSE}
if(!require("readr")) {
  install.packages("readr")
  library("readr")
}
if(!require("here")) {
  install.packages("here")
  library("here")
}
## set data dir
datadir <- here("data")
```

# User inputs

We begin by supplying values for the following parameters, which we use for trimming and lagging the covariates to the appropriate years.

```{r get_user_inputs}
## first & last years of fish data
yr_frst <- 1978
yr_last <- 2018

## min & max adult ages (years)
age_min <- 3
age_max <- 8

## time lags (years) for covariates
flow_lag <- 1
marine_lag <- 2
hrel_lag <- 2

## number of years for run forecasts
n_fore <- 0
```

# Retrieve covariates

Our analysis investigates 4 covariates as possible drivers of the population's instrinic growth rate:

1. Maximum river discharge in winter;
2. Minimum river discharge in summer;
3. North Pacific Gyre Oscillation;
4. Releases of hatchery-born juveniles.

## River discharge

We begin by getting the daily flow data from the US Geological Service [National Water Information System](http://waterdata.usgs.gov/nwis). We will use the direct link to the gage data from the Skagit River near Mount Vernon, WA (#12178100), beginning with the first year of fish data.

```{r get_flow_url}
## flow gage ID
flow_site <- 12178000
## get URL for flow data from USGS
flow_url <- paste0("https://waterdata.usgs.gov/nwis/dv",
                   "?cb_00060=on",
                   "&format=rdb",
                   "&site_no=",flow_site,
                   "&begin_date=",yr_frst,"-01-01",
                   "&end_date=",yr_last,"-12-31")
```

Next we retrieve the raw data file and print its metadata.

```{r get_flow_metadata, cache=TRUE}
## raw flow data from USGS
flow_raw <- read_lines(flow_url)
## lines with metadata
hdr_flow <- which(lapply(flow_raw, grep, pattern = "\\#")==1, arr.ind = TRUE)
## print flow metadata
print(flow_raw[hdr_flow], quote = FALSE)
```

Lastly, we extract the actual flow data for the years of interest and inspect the file contents.

```{r get_flows, cache=TRUE}
## flow data for years of interest
dat_flow <-  read_tsv(flow_url,
                      col_names = FALSE,
                      col_types = "ciDdc",
                      skip = max(hdr_flow)+2)
colnames(dat_flow) <- unlist(strsplit(tolower(flow_raw[max(hdr_flow)+1]),
                                      split = "\\s+"))
head(dat_flow)
```

We only need the 3rd and 4th columns, which contain the date (`datetime`) and daily flow measurements (``r grep("[0-9]$",colnames(dat_flow), value=TRUE)``). We will rename them to `date` and `flow`, respectively, and convert the flow units from "cubic feet per second" to "cubic meters per second".

```{r trim_dat_flow, cache=TRUE}
## keep only relevant columns
dat_flow <- dat_flow[c("datetime", grep("[0-9]$", colnames(dat_flow), value = TRUE))]
## nicer column names
colnames(dat_flow) <- c("date","flow")
## convert cubic feet to cubic meters
dat_flow$flow <- dat_flow$flow / 35.3147
## flow by year & month
dat_flow$year <- as.integer(format(dat_flow$date,"%Y"))
dat_flow$month <- as.integer(format(dat_flow$date,"%m"))
dat_flow <- dat_flow[,c("year","month","flow")]
```

### Winter maximum

We are interested in the maximum of the daily peak flows from October through March during the first year that juveniles are rearing in streams. This means we need to combine flow values from the fall of year $t$ with those in the winter and spring of year $t+1$. We also need to shift the flow data forward by `r flow_lag` year so they align with the juvenile life stage. Therefore, the flow time series will begin in `r yr_frst` and end in `r yr_last - age_min + n_fore + flow_lag`.

```{r wtr_flow, cache=TRUE}
## autumn flows in year t
flow_aut <- subset(dat_flow, (month>=10 & month<=12)
                   & year >= yr_frst & year <= yr_last-age_min+n_fore)
## spring flows in year t+1
flow_spr <- subset(dat_flow,
                   (month>=1 & month<=5)
                   & year >= yr_frst+flow_lag
                   & year <= yr_last-age_min+n_fore+flow_lag)
## change spr year index to match aut
flow_spr[,"year"] <- flow_spr[,"year"] - flow_lag
## combined flows indexed to brood year & calculate max flow
dat_flow_wtr <- aggregate(flow ~ year, data = rbind(flow_aut,flow_spr), max)
dat_flow_wtr[,"flow"] <- round(dat_flow_wtr[,"flow"], 1) 
## change year index to brood year
dat_flow_wtr[,"year"] <- dat_flow_wtr[,"year"] 
## for plotting purpose later
colnames(dat_flow_wtr)[2] <- "flow_wtr"
```

### Summer minimum

Retrieving the minimum flow juveniles would experience during their first summer (June through September) is straightforward.

```{r sum_flow, cache=TRUE}
## summer flows in year t
flow_sum <- subset(dat_flow, (month>=6 & month<=9)
                   & year >= yr_frst+flow_lag
                   & year <= yr_last-age_min+n_fore+flow_lag)
## change year index to brood year
flow_sum[,"year"] <- flow_sum[,"year"] - flow_lag
## combined flows indexed to brood year & calculate min flow
dat_flow_sum <- aggregate(flow ~ year, data = flow_sum, min)
dat_flow_sum <- round(dat_flow_sum, 2)
## for plotting purpose later
colnames(dat_flow_sum)[2] <- "flow_sum"
```

## North Pacific Gyre Oscillation

We used the monthly NPGO data provided by Emanuele Di Lorenzo of the Georgia Institute of Technology, which are available [here](http://www.o3d.org/npgo/npgo.php). We begin by downloading the raw NPGO data and viewing the metadata.

```{r get_NPGO_metadata, cache=TRUE}
## URL for NPGO data
url_NPGO <- "http://www.o3d.org/npgo/npgo.php"
## raw NPGO data 
NPGO_raw <- read_lines(url_NPGO)
## line with data headers
hdr_NPGO <- which(lapply(NPGO_raw,grep,pattern="YEAR")==1, arr.ind = TRUE)
## print PDO metadata
print(NPGO_raw[seq(hdr_NPGO)],quote = FALSE)
```

Next, we extract the actual NPGO indices for the years of interest and inspect the file contents. We also want the average NPGO annual index from January 1 through December 31 during the first year that the juvenile steelhead are in the ocean (i.e., during their second year of life). Therefore, we need NPGO values from `yr_frst + marine_lag == `r yr_frst+marine_lag`` through `yr_last - age_min + n_fore + marine_lag == `r yr_last - age_min + n_fore + marine_lag``.

```{r get_NPGO, cache=TRUE}
## number of years of data
n_yrs <- yr_last - yr_frst + 1
## NPGO data for years of interest
dat_NPGO <- read_table(url_NPGO, col_names = FALSE,
                       skip = hdr_NPGO + (yr_frst-1950)*12,
                       n_max = (n_yrs-1)*12)
colnames(dat_NPGO) <- c("year","month","NPGO")
## select only years of interest indexed by brood year 
dat_NPGO <- dat_NPGO[dat_NPGO$year >= yr_frst+marine_lag &
                     dat_NPGO$year <= yr_last-age_min+n_fore+marine_lag,]
dat_NPGO <- aggregate(dat_NPGO$NPGO, by = list(year = dat_NPGO$year), mean)
dat_NPGO <- data.frame(year = seq(yr_frst,yr_last-age_min+n_fore),
                       NPGO = dat_NPGO[,2])
dat_NPGO[,"NPGO"] <- round(dat_NPGO[,"NPGO"], 2)
```

## Hatchery releases

The numbers of hatchery fish released each year are provided in `/data/skagit_sthd_hrel.csv`. They have already been lagged 2 years (i.e., brood year + 2) to account for the potential competitive interactions during their juvenile life stage.

```{r get_hatchery_releases}
## get releases
dat_hrel <- read_csv(file.path(datadir, "skagit_sthd_hrel.csv"))
dat_hrel <- subset(dat_hrel, dat_hrel$year <= max(dat_NPGO$year))
```

# Archive covariates

The last thing we will do is combine the covariates into one data frame and write them to a file for use in the analysis.

```{r combine_covars, cache=TRUE}
## combine covariates
dat_cvrs <- Reduce(function(...) merge(..., all = TRUE),
                   list(dat_flow_wtr,
                        dat_flow_sum,
                        dat_NPGO,
                        dat_hrel))
## check table of covariates
head(dat_cvrs)
## write covariates to a file
write_csv(dat_cvrs, file.path(datadir, "skagit_sthd_covars_M.csv"))
```

